### 2016-07-06 Introduction.tex
a lot of changes

\section{Introduction}
The Standard Model of particle physics is a quantum field theory which relies on the principle of gauge invariance. It describes the world's fundamental particles and the interactions among them. It predicts the most accurately measured observables in physics and is therefore considered to be a very successful theory.\\
Besides these triumphs, it is known to be incomplete for it does, for example, not include dark matter. Because of these and other flaws, one studies extensions of the Standard Model. One mathematically very appealing extension is the concept of supersymmetry. However, besides the provision of a dark matter candidate and elegant solutions of, e.g. the hierarchy problem, supersymmetry does not come without problems. First of all, supersymmetric particles have not been discovered so far. Secondly, there are flaws like the ``SUSY flavor problem''. A non-trivial extension of supersymmetry - $R$-symmetry - may explain or solve these issues. Imposing supersymmetry and $R$-symmetry on the Standard Model, gives - according to the Haag-Lopuszanski-Sohnius theorem - the maximal symmetry group a reasonable quantum field theory can have.\\
In this thesis a minimal extension of the Standard Model including supersymmetry and $R$-symmetry - the MRSSM - will be considered. More specifically, the production cross section of strongly interacting supersymmetric particles is studied. It will turn out that there are significant changes in the production of squarks and gluinos when turning from the Minimal Supersymmetric Standard Model (MSSM) to the Minimal $R$-symmetric Supersymmetric Standard Model (MRSSM).\\
This thesis starts with summarizing the main aspects of the Standard Model before introducing the concept of supersymmetry and $R$-symmetry. Hereby, the relevant models including the minimal extension of the respective symmetry, the MSSM and the MRSSM are introduced. Section \ref{sec:SquarkGluinoTree} compares the production cross section of squark and gluinos in both models at tree-level. It will turn out that the most striking difference manifests in the channel of squark production. Therefore, the rest of the thesis focuses on the next-to-leading order calculation of this process: Section \ref{sec:VirtRealCorr} discusses the golden thread of the calculation of a next-to-leading order cross section including massless external particles before section \ref{sec:renMRSSM} gives detailed information about the renormalization of the MRSSM. This includes the explicit form of the required renormalization constants, reaching form the on-shell scheme over a mixture of $\overline{\mathrm{MS}}$- and zero-momentum subtraction scheme to supersymmetry restoring counterterms. This section also discusses what the most copious contributions to the renormalized matrix element are. The thesis closes with the representation of the next-to-leading order cross section for squark production in the form of $K$-factors.

>>>>

\section{Introduction}
The Standard Model of particle physics is a quantum field theory which relies on the principle of gauge invariance. It describes the world's fundamental particles and the interactions among them and predicts the most accurately measured observables in physics.\\
Besides these triumphs, it is known to be incomplete for it does, for example, not include dark matter or a description of gravity. Because of these and other flaws, one studies extensions of the Standard Model. One mathematically very appealing extension is the concept of supersymmetry. However, besides the provision of a dark matter candidate and elegant solutions of, e.g. the hierarchy problem, supersymmetry does not come without problems. Firstly, supersymmetric particles have not been discovered so far. Secondly, there are flaws like the ``SUSY flavor problem''. $R$-symmetry, a non-trivial extension of supersymmetry, may explain or solve these issues.
Another motivation to study $R$-symmetry is that, according to the Haag-Lopuszanski-Sohnius theorem, the combination of Poincaré symmetry, supersymmetry and $R$-symmetry is the maximal symmetry group a reasonable quantum field theory can have.\\
\ignore{Imposing supersymmetry and $R$-symmetry on the Standard Model, gives - according to the Haag-Lopuszanski-Sohnius theorem - the maximal symmetry group a reasonable quantum field theory can have.\\}
In this thesis a minimal extension of the Standard Model including supersymmetry and \mbox{$R$-symmetry} - the MRSSM - will be considered. More specifically, the production cross section of strongly interacting supersymmetric particles is studied. It will turn out that there are significant changes in the production of squarks and gluinos when turning from the Minimal Supersymmetric Standard Model (MSSM) to the Minimal $R$-symmetric Supersymmetric Standard Model (MRSSM).\\
This thesis starts with summarizing the main aspects of the Standard Model before introducing the concept of supersymmetry and $R$-symmetry. Hereby, the relevant models including the minimal extension associated to the respective symmetry, the MSSM and the MRSSM, are introduced. Section \ref{sec:SquarkGluinoTree} compares the production cross section of squark and gluinos in both models at \mbox{tree-level}. It will turn out that the most interesting difference manifests in the channel of squark production. Therefore, the rest of the thesis focuses on the next-to-leading order calculation of this process: Section \ref{sec:VirtRealCorr} discusses the golden thread of the calculation of a next-to-leading order cross section including massless external particles before section \ref{sec:renMRSSM} gives detailed information about the renormalization of the MRSSM. This includes the explicit form of the required renormalization constants, reaching form the on-shell scheme over a mixture of $\overline{\mathrm{MS}}$- and zero-momentum subtraction scheme to supersymmetry restoring counterterms. This section also discusses what the most copious contributions to the renormalized matrix element are. The thesis closes with the representation of the next-to-leading order cross section for squark production in the form of $K$-factors.

### 2016-07-06 NLO.tex
a lot of changes

It has been checked that for multiple different and quite distinct choices of the parameters $m_{\tilde{q}}$ and $m_{\tilde{g}}$ the double as well as the single poles of the virtual and real correction cancel up to machine precision. As the real corrections have been calculated analytically it is possible to quote the double pole. With

@@

Comment on how real corrections are processed.\\

>>>>

It has been checked that for multiple different and quite distinct choices of the parameters $m_{\tilde{q}}$ and $m_{\tilde{g}}$, the double as well as the single poles of the virtual and real correction cancel up to machine precision. As the real corrections have been calculated analytically it is possible to quote the double pole. With

@@

The computation of the real correction have been done by Wojciech Kotlarski, another member of the group. Comment on how real corrections are processed.\\

### 2016-07-06 RSymmetry.tex
TreeLevel SquarkProduction with errors including ALL channels redone
a lot of changes

\subsection{$R$-Symmetry Transformation}
For a $N=1$ supersymmetric theory, $R$-symmetry is a global $U(1)$ symmetry. It must not be confused with $R$-parity which is a discrete $\mathbb{Z}_2$ symmetry. A continuous global symmetry implies according to Noether's theorem a conserved charge. In the case of $R$-symmetry it is called $R$-charge and one therefore refers to $R$-symmetry as $U_R(1)$.\\
The defining property of $U_R(1)$ is that the anticommuting coordinates $\theta^\alpha$ and $\overline{\theta}^{\dot{\alpha}}$ transform like
\begin{align}
&\theta \to \mathrm{e}^{i\tau}\theta, && \overline{\theta} \to \mathrm{e}^{-i\tau}\overline{\theta},
\end{align}
where $\tau$ parametrizes the transformation. This in turn implies that the supersymmetry generators (see eq. \ref{{eq:SUSYGen}}) transform like
\begin{align}
& Q \to \mathrm{e}^{-i\tau} Q && \overline{Q} \to \mathrm{e}^{i\tau}\overline{Q}
\end{align}
and $R$-symmetry does not commute with supersymmetry 
\begin{align}
\left[ R,Q \right] = -Q && \left[ R,\overline{Q} \right] = \overline{Q},
\end{align}
meaning that superpartners do not have the same $R$-charge. The transformation of chiral and vector superfields reads
\begin{align}
& \hat{\Phi}(x,\theta,\overline{\theta}) \to \mathrm{e}^{ir_{\hat{\Phi}}\tau}\ \hat{\Phi}(x,\mathrm{e}^{i\tau}\theta,\mathrm{e}^{-i\tau}\overline{\theta}),\nonumber\\
&\hat{V}(x,\theta,\overline{\theta}) \to  \hat{V}(x,\mathrm{e}^{i\tau}\theta,\mathrm{e}^{-i\tau}\overline{\theta}),
\end{align}
i.e. the chiral superfields transform with arbitrary $R$-charge where the $R$-charge of vector superfield is restricted to be zero by the condition that they are real. If one inserts the component decomposition eq. \eqref{eq:superfielddecomp} of the superfields one can read off the $R$-charges of the component fields.
\begin{table}
\begin{center}
\begin{tabular}{c|c?c|c?c|c}
\multicolumn{2}{c?}{Superfield} & \multicolumn{2}{c?}{Boson} & \multicolumn{2}{c}{Fermion} \\
\hlinewd{2pt}
$\hat{\Phi}$ & $r_{\hat{\Phi}}$ & $A$ & $r_{\hat{\Phi}}$ & $\psi$ & $r_{\hat{\Phi}}-1$\\
$\hat{V}$ & 0 & $v^\mu$ & 0 & $\lambda$ & $+1$
\end{tabular}
\caption{This table shows the R-charges of a generic chiral and vector superfield.}
\end{center}
\end{table}

@@

\subsection{The Minimal $R$-Symmetric Supersymmetric Standard Model}
If one imposes $R$-symmetry upon the MSSM one is faced with a certain arbitrariness, i.e. the choice of the $R$-charges of the chiral superfields. A minimal $R$-symmetric extension of the MSSM is the minimal $R$-symmetric supersymmetric Standard Model (MRSSM)\cite{Kribs:2007ac}. In this model the $R$-charges are chosen in such a way, that every Standard Model particle has $R$-charge zero. Following this one obtains the $R$-charges of all particles which are summed up in table \ref{tab:R_charges_MRSSM}\cite{Diessner:2015bna}.
\begin{table}[H]
\begin{center}
\begin{tabular}{c?c|c?c|c?c|c}
Field & \multicolumn{2}{c?}{Superfield} & \multicolumn{2}{c?}{Boson} & \multicolumn{2}{c}{Fermion} \\
\hlinewd{2pt}
gauge Vector & $\hat{V}_s^a$, $\hat{V}_w^a$, $\hat{V}_Y$ & 0 & $G^a_\mu$, $W^a_\mu$, $B_\mu$ & 0 & $\lambda_s^a$, $\lambda_w^a$, $\lambda_Y$ & $+1$\\
matter & $\hat{L}$, $\hat{E}$ & $+1$ & $\tilde{l}_l$, $\tilde{e}^\dagger_R$ & $+1$ & $l_L$, $e_R$ & 0\\
 & $\hat{Q}$, $\hat{D}$, $\hat{U}$ & $+1$ & $\tilde{q}_L$, $\tilde{d}^\dagger_R$, $\tilde{u}^\dagger_R$ & $+1$ & $q_L$, $d_R$, $u_R$ & 0\\
$H$-Higgs & $\hat{H}_{d,u}$ & 0 & $H_{d,u}$ & 0 & $\tilde{H}_{d,u}$ & $-1$\\
\hline
$R$-Higgs & $\hat{R}_{d,u}$ & $+2$ & $R_{d,u}$ & $+2$ & $\tilde{R}_{d,u}$ & $+1$\\
adjoint chiral & $\hat{O}$, $\hat{T}$, $\hat{S}$ & 0 & $\sigma^a$, $\omega^a$, $\rho$ & 0 & $\chi_s^a$, $\chi_w^a$, $\chi_Y$ & $-1$
\end{tabular}
\caption{This table lists the $R$-charges of all superfields and their components in the MRSSM. The fields of the $R$-Higgs and the adjoint chiral superfields are not present in the MSSM}\label{tab:R_charges_MRSSM}
\end{center}
\end{table}

@@

& -\frac{m_{\tilde{q}}^2}{2}(|\tilde{q}_L|^2 + |\tilde{q}_R|^2) -m_{\sigma}^2\left|\sigma^a\right|^2  - m_g \overline{\tilde{g}}^a\tilde{g}^a\nonumber\\
& -\frac{1}{2}\left( g_s \tilde{q}_L^\dagger T^a \tilde{q}_L - g_s \tilde{q}_R T^{\ast a} \tilde{q}_R^\dagger - ig_s\sigma^{\dagger b}f^{abc}\sigma^c + \sqrt{2}m_g(\sigma^a + \sigma^{\dagger a}) \right)^2\label{eq:L_RSQCD}

@@

Observe that there is no 3 sgluon vertex, because of the antisymmetry of the structure constants $f^{abc}$. As already alluded to there are two distinct sgluons with different masses. The mass of the pseudoscalar $\sigma_0^a$ is given by the soft breaking parameter $m_{\sigma_0} = m_\sigma$ whereas the mass of the scalar $\phi^0$ is given by 

@@

The Feynman rules infered from this Lagrangian are given in Appendix \ref{sec:FeynmanRules}.

>>>>

\subsection{$R$-Symmetry Transformations}
For a $N=1$ supersymmetric theory, $R$-symmetry is a global $U(1)$ symmetry. It must not be confused with $R$-parity which is a discrete $\mathbb{Z}_2$ symmetry. A continuous global symmetry implies according to Noether's theorem a conserved charge. In the case of $R$-symmetry it is called $R$-charge and one therefore refers to $R$-symmetry as $U_R(1)$.\\
The defining property of $U_R(1)$ is that the anticommuting coordinates $\theta^\alpha$ and $\overline{\theta}^{\dot{\alpha}}$ transform like
\begin{align}
&\theta \to \mathrm{e}^{i\tau}\theta, && \overline{\theta} \to \mathrm{e}^{-i\tau}\overline{\theta},
\end{align}
where $\tau$ parametrizes the transformation. This in turn implies that the supersymmetry generators (see eq. \eqref{eq:SUSYGen}) transform like
\begin{align}
& Q \to \mathrm{e}^{-i\tau} Q && \overline{Q} \to \mathrm{e}^{i\tau}\overline{Q}
\end{align}
and $R$-symmetry does not commute with supersymmetry 
\begin{align}
\left[ R,Q \right] = -Q && \left[ R,\overline{Q} \right] = \overline{Q},
\end{align}
meaning that superpartners do not have the same $R$-charge. The transformation of chiral and vector superfields reads
\begin{align}
& \hat{\Phi}(x,\theta,\overline{\theta}) \to \mathrm{e}^{ir_{\hat{\Phi}}\tau}\ \hat{\Phi}(x,\mathrm{e}^{i\tau}\theta,\mathrm{e}^{-i\tau}\overline{\theta}),\nonumber\\
&\hat{V}(x,\theta,\overline{\theta}) \to  \hat{V}(x,\mathrm{e}^{i\tau}\theta,\mathrm{e}^{-i\tau}\overline{\theta}),
\end{align}
i.e. the chiral superfields transform with arbitrary $R$-charge where the $R$-charge of vector superfield is restricted to be zero by the condition that they are real. If one inserts the component decomposition eq. \eqref{eq:superfielddecomp} of the superfields one can read off the $R$-charges of the component fields.
\begin{table}
\begin{center}
\begin{tabular}{c|c?c|c?c|c}
\multicolumn{2}{c?}{Superfield} & \multicolumn{2}{c?}{Boson} & \multicolumn{2}{c}{Fermion} \\
\hlinewd{2pt}
$\hat{\Phi}$ & $r_{\hat{\Phi}}$ & $A$ & $r_{\hat{\Phi}}$ & $\psi$ & $r_{\hat{\Phi}}-1$\\
$\hat{V}$ & 0 & $v^\mu$ & 0 & $\lambda$ & $+1$
\end{tabular}
\caption{This table shows the R-charges of a generic chiral and vector superfield.}
\end{center}
\end{table}

@@

\subsection{The Minimal $R$-Symmetric Supersymmetric Standard Model}
If one imposes $R$-symmetry upon the MSSM one is faced with a certain arbitrariness, i.e. the choice of the $R$-charges of the chiral superfields. A minimal $R$-symmetric extension of the MSSM is the minimal $R$-symmetric supersymmetric Standard Model (MRSSM)\cite{Kribs:2007ac}. In this model the $R$-charges are chosen in such a way, that every Standard Model particle has $R$-charge zero. Following this one obtains the $R$-charges of all particles which are summed up in table \ref{tab:R_charges_MRSSM}\cite{Diessner:2014ksa}.
\begin{table}[H]
\begin{center}
\begin{tabular}{c?c|c?c|c?c|c}
Field & \multicolumn{2}{c?}{Superfield} & \multicolumn{2}{c?}{Boson} & \multicolumn{2}{c}{Fermion} \\
\hlinewd{2pt}
gauge Vector & $\hat{V}_s^a$, $\hat{V}_w^a$, $\hat{V}_Y$ & 0 & $G^a_\mu$, $W^a_\mu$, $B_\mu$ & 0 & $\lambda_s^a$, $\lambda_w^a$, $\lambda_Y$ & $+1$\\
matter & $\hat{L}$, $\hat{E}$ & $+1$ & $\tilde{l}_l$, $\tilde{e}^\dagger_R$ & $+1$ & $l_L$, $e_R$ & 0\\
 & $\hat{Q}$, $\hat{D}$, $\hat{U}$ & $+1$ & $\tilde{q}_L$, $\tilde{d}^\dagger_R$, $\tilde{u}^\dagger_R$ & $+1$ & $q_L$, $d_R$, $u_R$ & 0\\
$H$-Higgs & $\hat{H}_{d,u}$ & 0 & $H_{d,u}$ & 0 & $\tilde{H}_{d,u}$ & $-1$\\
\hline
$R$-Higgs & $\hat{R}_{d,u}$ & $+2$ & $R_{d,u}$ & $+2$ & $\tilde{R}_{d,u}$ & $+1$\\
adjoint chiral & $\hat{O}$, $\hat{T}$, $\hat{S}$ & 0 & $\sigma^a$, $\omega^a$, $\rho$ & 0 & $\chi_s^a$, $\chi_w^a$, $\chi_Y$ & $-1$
\end{tabular}
\caption{This table lists the $R$-charges of all superfields and their components in the MRSSM. The fields of the $R$-Higgs and the adjoint chiral superfields are not present in the MSSM}\label{tab:R_charges_MRSSM}
\end{center}
\end{table}

@@

& -\frac{m_{\tilde{q}}^2}{2}(|\tilde{q}_L|^2 + |\tilde{q}_R|^2) -m_{\sigma}^2\left|\sigma^a\right|^2  - m_{\tilde{g}} \overline{\tilde{g}}^a\tilde{g}^a\nonumber\\
& -\frac{1}{2}\left( g_s \tilde{q}_L^\dagger T^a \tilde{q}_L - g_s \tilde{q}_R T^{\ast a} \tilde{q}_R^\dagger - ig_s\sigma^{\dagger b}f^{abc}\sigma^c + \sqrt{2}m_{\tilde{g}}(\sigma^a + \sigma^{\dagger a}) \right)^2\label{eq:L_RSQCD}

@@

Observe that there is no 3 sgluon vertex, because of the antisymmetry of the structure constants $f^{abc}$. As already alluded to there are two distinct sgluons with different masses. The mass of the pseudoscalar $\sigma_0^a$ is given by the soft breaking parameter $m_{\sigma_0} = m_\sigma$ whereas the mass of the scalar $\phi_0$ is given by 

@@

The Feynman rules inferred from this Lagrangian are given in Appendix \ref{sec:FeynmanRules}.
